---
title: "Utano Market Survey Report"
author: "UlwaziHub"
output: pdf_document
editor: visual
---

## Introduction

This report presents the latest findings from the Utano Market Survey, conducted by UlwaziHUB Consulting Services in South Africa. The survey, which targeted a diverse population, was administered online between May 30 and June 5. A total of 391 participants responded, providing valuable insights into current market trends, healthcare access, and service preferences. The data collected serves as a critical foundation for understanding the needs and expectations of communities, and will inform the design and delivery of more inclusive and responsive health-related services.

## Methods

The Utano Market Survey employed a cross-sectional study design and was conducted entirely online using a structured digital questionnaire. The survey instrument was developed in English and translated into additional local languages to improve accessibility and comprehension. A combination of purposive and snowball sampling techniques was used to recruit participants through social media platforms, community networks, and digital outreach by local partners.

The questionnaire included both closed- and open-ended questions covering themes such as demographics, health-seeking behavior, service delivery preferences, and perceived barriers to healthcare. Data collection was anonymous, and participants provided informed consent electronically before beginning the survey. All data were securely stored and analyzed using statistical software to produce descriptive and inferential insights, including cross-tabulations and chi-square tests for associations between key variables such as gender and nationality.

## Results

This section presents the demographics and distribution of the results

```{r echo=FALSE,message=FALSE,warning=FALSE}
  library(googlesheets4)
  library(tidyverse)
  library(googledrive)
  library(summarytools)
  library(janitor)
  library(here)
  library(readxl)
  library(officer)
  library(flextable)

```
## Demographic Charecteristic of the Respondents

A crosstabulation of Nationality and Gender was performed to examine the distribution of respondents across these two demographic characteristics. This analysis provides valuable insights into the composition and representativeness of the study population. By assessing how gender is distributed within each nationality group, we gain a deeper understanding of the population dynamics and potential sociocultural patterns influencing participation or response behaviors.

```{r echo=FALSE ,warning=FALSE}

raw_data <-readxl::read_xlsx(here("Data","utano_raw_data.xlsx"),sheet = "Sheet1")

```

```{r echo=FALSE ,warning=FALSE}

raw_data %>%
  filter(gender %in% c("male", "female")) %>%
  mutate(nationality=gsub("_"," ",nationality)) %>% 
  mutate(nationality=str_to_sentence(nationality)) %>% 
  rename(Nationality=nationality) %>% 
  tabyl(Nationality, gender) %>%
    arrange(desc(male)) %>% 
  adorn_totals(where = "row") %>%             # Optional: add total row
  adorn_percentages("row") %>%                # Calculate row percentages
  adorn_pct_formatting(digits = 0) %>%        # Format percentages
  adorn_ns(position = "front") %>%  
    rename(
    `Male (n %)` = male,
    `Female (n %)` = female
  ) %>% 

  flextable() %>%
  autofit()



```

A chi-square test of independence was conducted to examine the relationship between gender and nationality among participants. The analysis revealed a statistically significant association between gender and nationality, χ²(df = r chisq_result$parameter) = r round(chisq_result$statistic, 2), p = r p_val. This suggests that the distribution of gender differs significantly across nationalities in the dataset. The observed differences were supported by the cross-tabulated counts and row-wise percentages, with variations noted particularly in the proportions of males and females within each nationality group.

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Run chi-square test
chi_data <- raw_data %>%
  filter(gender %in% c("male", "female")) %>%
  select(nationality, gender)

chisq_result <- chisq.test(table(chi_data$nationality, chi_data$gender))

# Extract and format p-value
p_val <- signif(chisq_result$p.value, 3)
p_text <- paste("Chi-square p =", p_val)


# Display

p_text
```
## Healthcare Needs & Usage.
Understanding how different genders engage with various service types can provide valuable insights into consumer preferences and accessibility. This analysis explores the distribution of male and female preferences across different service types.The chart highlights a clear gender difference in service preferences. Females predominantly chose Gold, Platinum, and Silver services, with Gold being the most favored. In contrast, males showed a stronger preference for Lithium services, demonstrating a noticeable skew toward this option. This pattern suggests potential gender-driven trends that could inform service customization and marketing strategies
```{r echo=FALSE,warning=FALSE,message=FALSE}
raw_data %>% filter(gender %in% c("male","female"),service_type!="NA") %>% 
  mutate(gender=str_to_sentence(gender))%>% 
                      ggplot() +
  aes(x = service_type, fill = gender) +
  geom_bar(position = "dodge") +
  scale_fill_hue(direction = 1) +
  theme_linedraw() +
  theme(legend.justification = "right")+
   labs(x = "Service Type", y = "Frequency",fill="Gender")  # Add axis titles here
  

```
Understanding how different Nationality engage with various service types can provide valuable insights into consumer preferences and accessibility.The chart below highlights how employment status influence service preferences. Those in the informal sector predominantly chose Gold  while those in the formal sector,show stronger preference in Lithium, Platinum and Silver services. 
```{r echo=FALSE,warning=FALSE,message=FALSE}
raw_data %>% filter(gender %in% c("male","female"),occupation!="NA",service_type!="NA") %>% 
  mutate(occupation=as.factor(str_to_sentence(occupation))) %>% 
  mutate(occupation=fct_relevel(occupation, "informal","formal","ownbusiness","unemployed"))%>% 
                      ggplot() +
  aes(x = service_type, fill = occupation) +
  geom_bar(position = "dodge") +
  scale_fill_hue(direction = 1) +
  theme_linedraw() +
  theme(legend.justification = "right")+
   labs(x = "Service Type", y = "Frequency",fill="Gender")  # Add axis titles here
```

## SECTION C: Willingness to sign up for the product.
Language Preference and Choice of Healthcare Provider – Gender Disaggregation

The majority of both female and male respondents indicated that language plays a significant role in their choice of healthcare provider, with 163 females (50%) and 160 males (50%) agreeing with this statement. A smaller proportion disagreed—17 females (57%) vs. 13 males (43%)—suggesting that more females than males feel language is important, though the numbers remain relatively low. Among those who were uncertain, the response was balanced, with 6 females (46%) and 7 males (54%) indicating they didn’t know whether language influences their provider choice. Overall, 366 respondents participated in this assessment, with a slightly higher representation of females (51%) than males (49%).

This suggests that language is a key consideration for both genders when selecting a healthcare provider, highlighting the importance of language-concordant care in health service delivery (Table below)
```{r warning=FALSE,message=FALSE,echo=FALSE}
raw_data %>% filter(gender %in% c("male","female"),occupation!="NA",service_type!="NA",language_pref!="NA") %>% 
   mutate(language_pref=str_to_sentence(language_pref),gender=str_to_sentence(gender)) %>% 
  mutate(language_pref=ifelse(language_pref=="Dontknow","Dont know",language_pref))%>% 
  rename(Language=language_pref) %>% 
  tabyl(Language,gender) %>% 
  adorn_totals(where = "col") %>%  
   adorn_totals(where = "row") %>%             # Optional: add total row
  adorn_percentages("col") %>%                # Calculate row percentages
  adorn_pct_formatting(digits = 0) %>%        # Format percentages
  adorn_ns(position = "front") %>% 
  flextable() %>% 
  autofit()
```

```{r warning=FALSE,message=FALSE,echo=FALSE}
raw_data %>% filter(gender %in% c("male","female"),occupation!="NA",service_type!="NA",gender_pref!="NA") %>% 
   mutate(gender_pref=str_to_sentence(gender_pref),gender=str_to_sentence(gender)) %>% 
  mutate(gender_pref=ifelse(gender_pref=="Dontknow","Dont know",gender_pref))%>% 
  rename(Gender=gender_pref) %>% 
  tabyl(Gender,gender) %>% 
  adorn_totals(where = "col") %>%  
   adorn_totals(where = "row") %>%             # Optional: add total row
  adorn_percentages("col") %>%                # Calculate row percentages
  adorn_pct_formatting(digits = 0) %>%        # Format percentages
  adorn_ns(position = "front") %>% 
  flextable() %>% 
  autofit()
```

